---
title: "External Data Sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{External Data Sources}
  %\VignetteEngine{rmarkdown::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Test data for FastQDesign is hosted on Zenodo for reproducibility and easy access.

## Download Links

### BAM Files (scRNA-seq alignments)

Download from: [https://zenodo.org/records/19073177](https://zenodo.org/records/19073177)

Contains Cellranger-aligned BAM files with cell barcodes (CB tag) and UMI (UB tag):

| File | Size |
|------|------|
| `P1_AI_possorted_genome_bam.bam` | 13.1 GB |
| `P1_AI_possorted_genome_bam.bam.bai` | 4.1 MB |
| `P1_BM_possorted_genome_bam.bam` | 14.4 GB |
| `P1_BM_possorted_genome_bam.bam.bai` | 4.4 MB |

### Seurat Objects (processed data)

Download from: [https://zenodo.org/records/19072084](https://zenodo.org/records/19072084)

Contains pre-processed Seurat objects:

| File | Size |
|------|------|
| `bam_downsample_list.rds` | 77.9 MB |
| `reference_list.rds` | 191.2 MB |

## URL Streaming (No Download Required)

BAM files can be accessed directly from Zenodo URL without downloading.

### Using Rsamtools

```{r, eval = FALSE}
library(Rsamtools)

bam_url <- "https://zenodo.org/records/19073177/files/P1_AI_possorted_genome_bam.bam"
bai_url <- "https://zenodo.org/records/19073177/files/P1_AI_possorted_genome_bam.bam.bai"

bf <- BamFile(bam_url, index = bai_url)
header <- scanBamHeader(bf)
```

### Using fastF

fastF supports URL streaming for `extract` and `crb` subcommands:

```{r, eval = FALSE}
# Extract CR tags (UMI sequences)
# Processes 182M reads in ~8 minutes from URL
FastQDesign::fastF_extract(
  bam = "https://zenodo.org/records/19073177/files/P1_AI_possorted_genome_bam.bam",
  tag = "CR",
  type = 0
)
```

The `extract` subcommand streams BAM via htslib's HTTP support. For very large files, this is efficient as only necessary regions are downloaded.

## Local Download

```{r, eval = FALSE}
# Download BAM
download.file(
  "https://zenodo.org/records/19073177/files/P1_AI_possorted_genome_bam.bam",
  destfile = "P1_AI_possorted_genome_bam.bam"
)

# Download Seurat object
download.file(
  "https://zenodo.org/records/19072084/files/reference_list.rds",
  destfile = "reference_list.rds"
)

library(Seurat)
seurat_obj <- readRDS("reference_list.rds")
```

## Citation

```
Wang Y (2026). FastQDesign Test Data. Zenodo.
https://zenodo.org/records/19072084
https://zenodo.org/records/19073177
```
