---
title: "Experiment Design with FastQDesign"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experiment Design with FastQDesign}
  %\VignetteEngine{rmarkdown::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

FastQDesign helps investigators plan scRNA-seq experiments by optimizing the number of cells and sequencing depth given a budget constraint.

## Data Source

Power analysis data (`df_power`) is included with the package:

```r
# Load power analysis data from package
df_power <- read.csv(system.file("data", "AIBM_power.csv", package = "FastQDesign"))
```

## Basic Usage

```{r eval=FALSE}
library(FastQDesign)

# Load power analysis data
df_power <- read.csv(system.file("data", "AIBM_power.csv", package = "FastQDesign"))

# Define experiment parameters
budget <- 7500  # Total budget in USD
power_threshold <- 0.7  # Statistical power target
flowcell_capacities <- c(10^7, 5 * 10^7, 2 * 10^8)  # Reads per flowcell
flowcell_costs <- c(1000, 2000, 3000)  # Cost per flowcell
library_costs <- 5000  # Library preparation cost

# Run experiment design
rst <- FastQDesign(
  df_power = df_power,
  budget = budget,
  power_threshold = power_threshold,
  flowcell_capacities = flowcell_capacities,
  flowcell_costs = flowcell_costs,
  library_costs = library_costs,
  reads_valid_rate = 0.9
)
```

## Output

The `FastQDesign()` function returns a list containing:

- **$p_ind**: Individual sample design plot
- **$p_share**: Shared sample design plot  
- **$design**: Design results table

## Parameters

| Parameter | Description |
|-----------|-------------|
| `df_power` | DataFrame with n_cells, expected_reads_per_cell, power columns |
| `budget` | Total experiment budget |
| `power_threshold` | Target statistical power |
| `flowcell_capacities` | Vector of flowcell read capacities |
| `flowcell_costs` | Vector of flowcell costs |
| `library_costs` | Library preparation cost |
| `reads_valid_rate` | Expected valid reads rate |

## Plotting

Save the generated plots:

```{r eval=FALSE}
# Save individual plot
ggsave("design_individual.pdf", rst$p_ind, width = 8, height = 6)

# Save combined plot
combined <- patchwork::wrap_plots(rst$p_ind, rst$p_share, ncol = 1)
ggsave("design_combined.pdf", combined, width = 10, height = 10)
```
