---
title: "Using fastF Tool"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using fastF Tool}
  %\VignetteEngine{rmarkdown::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `fastF` tool is a C-based command-line utility for scRNA-seq experiment design. It provides several subcommands for processing FastQ and BAM files.

## Installation

The fastF binary is automatically installed with the FastQDesign package:

```{r eval=FALSE}
devtools::install_github("yuw444/FastQDesign")
```

## fastF Subcommands

The package includes test data for trying out fastF commands:

```{r eval=FALSE}
# Paths to package test data
test_R1 <- system.file("data", "test_R1.fastq.gz", package = "FastQDesign")
test_R2 <- system.file("data", "test_R2.fastq.gz", package = "FastQDesign")
test_bam <- system.file("data", "test.bam", package = "FastQDesign")
test_features <- system.file("data", "features.tsv.gz", package = "FastQDesign")
test_barcodes <- system.file("data", "barcodes.tsv.gz", package = "FastQDesign")
```

### 1. Find Cell Barcode Whitelist

Extract cell barcode whitelist from FastQ files:

```{r eval=FALSE}
# Find whitelist from R1 FastQ file
FastQDesign::fastF_freq(
  R1 = test_R1,
  out = "output_dir",
  len = 16L
)
```

This creates `whitelist.txt` in the output directory.

### 2. Filter FastQ Files

Filter FastQ files using a whitelist and downsampling:

```{r eval=FALSE}
FastQDesign::fastF_filter(
  R1 = test_R1,
  R2 = test_R2,
  whitelist = system.file("data", "whitelist.txt", package = "FastQDesign"),
  out = "filtered_dir",
  len = 16L,
  seed = 42L,
  rate = 0.1
)
```

### 3. Extract CR/CB Tags from BAM

Summarize cell barcode and UMI information:

```{r eval=FALSE}
FastQDesign::fastF_crb(
  bam = test_bam,
  out = "output.tsv.gz"
)
```

### 4. Convert BAM to Matrix

Convert BAM to sparse matrix format:

```{r eval=FALSE}
FastQDesign::fastF_bam2db(
  bam = test_bam,
  feature = test_features,
  barcode = test_barcodes,
  out = "matrix_dir"
)
```

### 5. Extract Custom Tags

Extract custom tags from BAM files:

```{r eval=FALSE}
FastQDesign::fastF_extract(
  bam = test_bam,
  tag = "CB"
)
```

## Running from Command Line

You can also run fastF directly from the command line:

```bash
# Find whitelist
fastF freq -R input_R1.fastq.gz -l 16 -o output_dir/

# Filter FastQ
fastF filter -R input_R1.fastq.gz -r input_R2.fastq.gz -w whitelist.txt -o filtered/

# Extract CR/CB tags
fastF crb -b input.bam -o output.tsv.gz
```
